Nathan Greenslit Data Analysis Project

### 2023-04-20 code/analysis_code/env/env_par.qmd
mansuscript edits

```{r}
don<-
  daily %>%
  ggplot() + geom_line(
  aes(
    x = date,
    y = don,
    color = site),
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "DON (µM)",
  #  title = "Temperature across Daily Time Series by Site",
    color = "Site")+
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
   axis.ticks.x = element_blank(),
    axis.title.y = element_text(size = 15),
    axis.text.y = element_text(size = 7),
    legend.position = "none",
    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
    scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))
```

@@

```{r}
doc<-
  daily %>%
  ggplot() + geom_line(
  aes(
    x = date,
    y = doc,
    color = site),
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "DOC (µM)",
  #  title = "Temperature across Daily Time Series by Site",
    color = "Site")+
  theme(
    axis.title.x = element_blank(),
   # axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(size = 15),
    axis.text.y = element_text(size = 7),
    legend.position = "none",
    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
    scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))
```

@@

```{r}
tdn<-
   daily %>%
  ggplot() + geom_line(
  aes(
    x = date,
    y = tdn,
    color = site),
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "TDN (µM)",
    color = "Site")+
  theme(
    plot.title= element_text(hjust = 0.5, size = 20),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
   # axis.text.x = element_blank(),
    axis.title.y = element_text(size = 15),
    axis.text.y = element_text(size = 7),
    legend.key.size = unit(0.3,'cm'),
    legend.position = "none")+
    scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))
```

@@

```{r}
chl<-
  daily %>%
  ggplot() + geom_line(
  aes(
    x = date,
    y = chl,
    color = site),
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "Chlorophyl (µg/L)",
    color = "Site")+
  theme(
    plot.title= element_text(hjust = 0.5, size = 20),
    axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
   # axis.text.x = element_blank(),
    axis.title.y = element_text(size = 15),
    axis.text.y = element_text(size = 7),
    legend.key.size = unit(0.3,'cm'),
    legend.position = "none")+
    scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))
```

@@

```{r, fig.width= 8, fig.height=4}
fig<- ggarrange(#temp,
                #sal,
               # do,
                #nn,
                #ortho,
                #don,
                #doc,
                #pH,
                #amm,
                tn,
                sil,
                cop,
                chl,
                secchi,
                din,
                din_dip,
                tdn,
                ncol = 3, 
                nrow = 3,
                align = "v")

>>>>

```{r}
don<-
  daily %>%
  ggplot() + geom_line(
  aes(
    x = date,
    y = don,
    color = site),
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "DON (µM)",
  #  title = "Temperature across Daily Time Series by Site",
    color = "Site")+
  theme(
    #axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
   #axis.ticks.x = element_blank(),
    axis.title.y = element_text(size = 15),
    axis.text.y = element_text(size = 7),
    legend.position = "none",
    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
    scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))
```

@@

```{r}
doc<-
  daily %>%
  ggplot() + geom_line(
  aes(
    x = date,
    y = doc,
    color = site),
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "DOC (µM)",
  #  title = "Temperature across Daily Time Series by Site",
    color = "Site")+
  theme(
    #axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    #axis.ticks.x = element_blank(),
    axis.title.y = element_text(size = 15),
    axis.text.y = element_text(size = 7),
    legend.position = "none",
    legend.background = element_rect(colour = 'black', fill = 'white', linetype='solid')) + #Moves legend onto plot 
    scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))
```

@@

```{r}
tdn<-
   daily %>%
  ggplot() + geom_line(
  aes(
    x = date,
    y = tdn,
    color = site),
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "TDN (µM)",
    color = "Site")+
  theme(
    plot.title= element_text(hjust = 0.5, size = 20),
   # axis.title.x = element_blank(),
    #axis.ticks.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y = element_text(size = 15),
    axis.text.y = element_text(size = 7),
    legend.key.size = unit(0.3,'cm'),
    legend.position = "none")+
    scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))
```

@@

```{r}
chl<-
  daily %>%
  ggplot() + geom_line(
  aes(
    x = date,
    y = chl,
    color = site),
  size = 1)+
  theme_bw()+
  labs(
    x = "Date",
    y = "Chlorophyl (µg/L)",
    color = "Site")+
  theme(
    plot.title= element_text(hjust = 0.5, size = 20),
    #axis.title.x = element_blank(),
   # axis.ticks.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.y = element_text(size = 15),
    axis.text.y = element_text(size = 7),
    legend.key.size = unit(0.3,'cm'),
    legend.position = "none")+
    scale_color_manual(values = c("Blind Oso" = "#809C52",
                                "Canals" = "#ba4716",
                                "Gulf" = "4c848c"))
```

@@

```{r, fig.width= 8, fig.height=4}
fig<- ggarrange(temp,
                sal,
                #do,
                nn,
                ortho,
                don,
                doc,
                #pH,
                #amm,
                #tn,
                #sil,
                #cop,
                tdn,
                #secchi,
                #din,
                #din_dip,
                chl,
                ncol = 4, 
                nrow = 2,
                align = "v")

### 2023-04-20 products/manuscript/Manuscript.qmd
mansuscript edits

![Figure 1: Total Daily Vibrio at All Sites](images/paste-C33B6E43.png)

@@

![Figure 2: Environmental parameters across daily time series](images/00000f%20(1).png)

<!--# Add detail on other env variables -->

@@

![Figure 4](images/000055.png){width="481"}

![Table 1: Univariate Linear Regression Model Metrics](images/Screen%20Shot%202023-04-19%20at%201.33.37%20PM.png){width="367"}

@@

Overall, *Vibrio* growth was observed following dust events at both the Gulf and Blind Oso. Similar to previous findings peaks in growth were observed 24-48 hours following dust deposition [@westrichVibrioPopulationDynamics2018a; @westrichSaharanDustNutrients2016a]. Linear regression analysis confirmed a correlation between growth and dust input (R2 = 0.49, 0.56, 0.44 for Blind Oso, Canals, and Gulf respectively).

@@

However, the environmental dynamics of each site most likely contributed the degree of response per site. Multivariate models were tested using specific environmental parameters: dust, DOM (DON, DOC), temperature, salinity, orthophosphate, and nitrate+nitrite. Ultimately, a model that included water chemistry, dust, nutrient, and dissolved organic matter parameters provided the best performance (R2 = 0.73, 0.7, and 0.76 for Blind Oso, Canals, and Gulf respectively), suggesting that while dust may contribute to a growth response in *Vibrio* following dust input, the site specific dynamics aid in dictating the degree of response observed.

>>>>

![Figure 1: Total Daily Vibrio at All Sites](images/paste-C33B6E43.png){width="1491"}

@@

![Figure 2:](images/000019.png)

@@

Linear regression models between dust AOD and copies per mL of Vibrio. R2 values indicates the strength of the linear model, and the Pearson Correlation Coefficient indicates the strength of the linear relationship between the two variables.

![Figure 4](images/000055.png){width="1491"}

![Table 1: Uni-variate Linear Regression Model Metrics](images/Screen%20Shot%202023-04-19%20at%201.33.37%20PM.png){width="367"}

@@

#### Vibrio and Dust

*Vibrio* growth was found to significantly correlate with dust aerosol optical density at Blind Oso (r = 0.7, p = 0.01), Canals (r = 0.75, p = 0.005), and the Gulf (r = 0.67, p = 0.02). Similar to previous findings peaks in growth were observed 24-48 hours following dust deposition, with both the Gulf and Blind Oso having a peak response 24hr following peak dust deposition, and the Canals 48hr after peak deposition [@westrichVibrioPopulationDynamics2018a; @westrichSaharanDustNutrients2016a]. <!--# Add xfold increases -->.

The introduction of dust constituents have the potential to serve as a temporary alleviation from the limitation of nutrients that are essential for growth. *Vibrio* species have one of the fastest doubling times among bacteria \[@gibsonDistributionBacterialDoubling2018 ; @eagonPSEUDOMONASNATRIEGENSMARINE1962\], and are able to rapidly exploit introduced nutrients due their rapid motility, diverse genomic repertoire, and iron-chelating complexes \[@ringgaardChemotaxisArraysVibrio2018 ; @okadaVibriosCommonlyPossess2005 ; @payneVibrioIronTransport2016\]. These advantageous traits have resulted in *Vibrio* being classified as 'oppurtunitrophs'. The following declines in *Vibrio* gene copies/mL may be attributed to topdown control, such as viral lysis <!--# Cite -->, or bottom-up controls due to an exhaustion of the resources provided by the dust.

@@

Multivariate models were tested using specific environmental parameters that have been associated with *Vibrio* population dynamics [@bullington2022]. Ultimately, a model that included water chemistry, dust, nutrient, and dissolved organic matter parameters provided the best performance (Table 3, Table 4, Table5), suggesting that while dust may contribute to a growth response in *Vibrio*, the site specific nutrient and chemistry dynamics play an important role in dictating the degree of response observed.

<!--# Add monthly average of nutrients -->

Similar to previous work done in coastal and offshore sites, the Gulf experienced

Blind Oso Bay is an enclosed shallow tributary that frequently expereince high levels of nutrients (total dissolved nitrogen) and chlorophyll due to its proximity to the Oso Wastewater Plant (OSP). The Bay experienced a dramatic drop in salinity from July 11th to July 19th. This decrease in salinity could be attributed to discharge from the OSP.
